Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2023-04-05 arxiv/main.tex

\subsection*{A closed-form expression for a unified Gompertz-Logistic model}
Finally, we are positioned to arrive at a closed-form expression for a finite number of interaction terms in this modified logistic model, which represents the infection term in the SIR model. Working with a single unitless growth variable and omitting the multivariate network without loss of generality, we start with

\begin{equation}
\frac{d \ln{x}}{dt} = \sum_{i=1}^N s^i/i.
\end{equation}
Using standard integrals we see that the multiple terms amount to a Gompertz growth term adjusted by a Hypergeometric function,
\begin{eqnarray*}
  \sum_{i=1}^N s^i/i &=& \int_s \sum_{i=0}^{N-1} s^i ds\\
  &=&\int_s \frac{1-s^N}{1-s} ds\\
  &=&- \frac{{}_{1}F_{2}({1,N+1,N+2;s})}{N+1}s^{N+1} - \ln(1-s),
\end{eqnarray*}
where ${}_1F_{2}$ is Gauss' Hypergeometric function, and integration limits are indefinite. In summary, the modified logistic function emerges as
 \begin{equation}
 \label{eq:modLogistic}
\frac{d}{dt}\ln{x(t)} = - \frac{{}_{1}F_{2}({1,N+1,N+2;1-x(t)})}{N+1}(1-x(t))^{N+1} - \ln{x(t)},
 \end{equation}
 which represents the novel bridge between logistic and Gompertzian growth. Note that for $N=1$, this relative growth rate becomes the standard logistic function, while increasing the number of higher order terms will move the equation closer to the Gompertz curve as the Hypergeometric term approaches zero. This modification to the logistic function does not have a 1-to-1 correspondence with the Richards model, but serves instead as an alternative when encoding non-linear or collaborative growth effects (see Fig. \ref{fig:hypergeometric}). This novel growth equation can thus be included in the SIR equation for the infected.

@@

We can now further appreciate the Richards parametrization as a transition from non-collaborative to maximally collaborative growth as $\theta \rightarrow 0$. 
This feature of $\theta$ was also obtained by Petroni et al. (2020) \cite{petroni2020logistic} by interpreting the $\theta-$logistic growth rate in Eq. \ref{eq:rypdalGamma} as non-linear resource availability dependent on the overall abundance, with linearity reached at $\theta=1$. The maximally coherent scenario implies that all units in the network are exposed to the stressor simultaneously.
Molski and Konarski (2003) \cite{molski2003coherent} made a similar observation that Gompertzian growth is the coherent state in a quantum mechanical system with a time-dependent potential, an interpretation which sheds further light on the temporal nature of the postulated stress response.
This quantum mechanical system has also been used to describe coherent energy states of diatomic molecules in space \cite{morse1929diatomic}. 
In the field of quantum physics, \emph{coherence} is a well-defined mathematical property first explored by Glauber (1963) \cite{glauber1963coherent} in the context of electromagnetic fields. The fact that we observe the Gompertz curve in both the microscopic quantum scales and the macroscopic national scales is noteworthy, and suggests that both systems share commonalities and means of communication.

>>>>

\subsection*{A closed-form expression for a unified Gompertz-Logistic model}
Finally, we are positioned to arrive at a closed-form expression for a finite number of interaction terms in this modified logistic model, which represents the infection term in the SIR model. Working with a single unitless growth variable and omitting the multivariate network without loss of generality, we start with the well-known identity

\begin{equation}
\frac{d \ln{x}}{dt} = \sum_{i=1}^N s^i/i.
\end{equation}
Using standard integrals we see that this sum amounts to a Gompertz growth term adjusted by a Hypergeometric function,
\begin{eqnarray*}
  \sum_{i=1}^N s^i/i &=& \int_s \sum_{i=0}^{N-1} s^i ds\\
  &=&\int_s \frac{1-s^N}{1-s} ds\\
  &=&- \frac{{}_{1}F_{2}({1,N+1,N+2;s})}{N+1}s^{N+1} - \ln(1-s),
\end{eqnarray*}
where ${}_1F_{2}$ is Gauss' Hypergeometric function, and integration limits are indefinite. In summary, the modified logistic function emerges as
 \begin{equation}
 \label{eq:modLogistic}
\frac{d}{dt}\ln{x(t)} = - \frac{{}_{1}F_{2}({1,N+1,N+2;1-x(t)})}{N+1}(1-x(t))^{N+1} - \ln{x(t)},
 \end{equation}
 which represents the novel bridge between logistic and Gompertzian growth. Note that for $N=1$, this relative growth rate becomes the standard logistic function, while increasing the number of higher order terms will move the equation closer to the Gompertz curve as the Hypergeometric term approaches zero. This modification to the logistic function does not have a 1-to-1 correspondence with the Richards model, but serves instead as an alternative when encoding non-linear or collaborative growth effects (see Fig. \ref{fig:hypergeometric}). This novel growth equation can thus be included in the SIR equation for the infected.

@@

We can now further appreciate the Richards parametrization as a transition from non-collaborative to maximally collaborative growth as $\theta \rightarrow 0$. 
This feature of $\theta$ was also obtained by Petroni et al. (2020) \cite{petroni2020logistic} by interpreting the $\theta-$logistic growth rate in Eq. \ref{eq:rypdalGamma} as non-linear resource availability dependent on the overall abundance, with linearity reached at $\theta=1$. Used in a mortality model, the maximally coherent scenario implies that all units in the network are exposed to the stressor simultaneously.
Molski and Konarski (2003) \cite{molski2003coherent} made a similar observation that Gompertzian growth is the coherent state in a quantum mechanical system with a time-dependent potential, an interpretation which sheds further light on the temporal nature of the postulated stress response.
This quantum mechanical system has also been used to describe coherent energy states of diatomic molecules in space \cite{morse1929diatomic}. 
In the field of quantum physics, \emph{coherence} is a well-defined mathematical property first explored by Glauber (1963) \cite{glauber1963coherent} in the context of electromagnetic fields. The fact that we observe the Gompertz curve in both the microscopic quantum scales and the macroscopic national scales is noteworthy, and suggests that both systems share commonalities and means of communication.
